---
title: "Discussion 3 Nov"
format: html
editor: visual
---

## TidyTuesday Dataset

<https://github.com/rfordatascience/tidytuesday/blob/main/data/2025/2025-10-28/readme.md>

```{r, warning=FALSE, message=FALSE}
library(dplyr)
library(readr)
# install.packages("tidytuesdayR")
prizes <- readr::read_csv("https://raw.githubusercontent.com/rfordatascience/tidytuesday/main/data/2025/2025-10-28/prizes.csv")

# tuesdata <- tidytuesdayR::tt_load('2025-10-28')
#tuesdata <- tidytuesdayR::tt_load(2025, week = 43)
prizes <- prizes %>%
  mutate(
    prize_year = as.numeric(prize_year)
  )
# prizes <- tuesdata$prizes
```

```{r}
length(unique(prizes$degree_institution))
```

```{r}
length(unique(prizes$prize_genre))
```

```{r}
length(unique(prizes$prize_alias))
```

```{r}
length(unique(prizes$prize_name))
```

```{r}
length(unique(prizes$prize_id))
```

```{r}
prizes |> count(uk_residence, person_role, ethnicity_macro)
```

```{r}
length(unique(prizes$ethnicity_macro))
```

```{r}
colnames(prizes)

```

Clean steps:

```{r}
library(tidyverse)
library(dplyr)

prizes_clean <- prizes %>%
  mutate(
    prize_genre = as.factor(prize_genre),
    prize_alias = as.factor(prize_alias),
    prize_name = as.factor(prize_name),
    person_role = as.factor(person_role),
    gender = as.factor(gender),
    sexuality = as.factor(sexuality),
    ethnicity = as.factor(ethnicity),
    ethnicity_macro = as.factor(ethnicity_macro),
    highest_degree = as.factor(highest_degree),
    degree_field = as.factor(degree_field),
    degree_field_category = as.factor(degree_field_category),
    uk_residence = as.logical(uk_residence)
  ) %>%

  mutate(
    across(where(is.character), \(x) str_squish(x))
  ) %>%

  mutate(
    prize_id = as.character(prize_id),
    person_id = as.character(person_id),
    book_id = as.character(book_id),
    viaf = as.character(viaf)
  ) %>%

  distinct()
write.csv(prizes_clean, "prizes_clean.csv", row.names = FALSE)
```

## Deliverable

-   Create a **Shiny app** with an interactive interface that allows users to easily explore the dataset using relevant filters.

-   Your app should include multiple **web pages (tabs)** that address key questions about the dataset. These questions can be based on those discussed in class or ones you develop on your own — bonus points will be awarded for original questions that demonstrate insight or creativity.

-   The app should include **data visualizations** that help communicate your findings clearly.

-   When submitting, please provide a **GitHub repository** containing:

    -   `app.R` (the main file to run the app)

    -   Any additional scripts used in the app

    -   The cleaned dataset

    -   A working version of the app that can be launched directly from `app.R`
